---
title: "Logistic_Analysis"
author: "Xiao Chen"
date: "2023-03-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## import data

```{r}
rm(list = ls())
df <- read.csv("Trajectory_Data_Clean_withGroupInfo.csv")

# select the right group weight number according to the grouping number
df$Group_Weights <- ifelse(df$X_traj_Group == 1, df$X_traj_ProbG1,
                          ifelse(df$X_traj_Group == 2, df$X_traj_ProbG2,
                                 ifelse(df$X_traj_Group == 3, df$X_traj_ProbG3,
                                        ifelse(df$X_traj_Group == 4, df$X_traj_ProbG4, NA)
                                        )))
```

## add a variable named "Anxiety" and "biological sex"

0: no reported comorbidities 1: with anxiety (may include other comorbidities) 2: only other comorbidities

```{r}
library(tidyverse)
library(dplyr)

df <- df %>%
  mutate(
    Anxiety = case_when(
      grepl("Anxiety", df$ComorbiditiesNotDepression)~1,
      TRUE~0
    )
  )
df$Anxiety <- as.factor(df$Anxiety)

# the observation who identify herself as non-binary (=5) is biologically female (=1)
df <- df %>%
  mutate(
    BiologicalSex = case_when(
      BiologicalSexandGenderifDi == 1~1,
      BiologicalSexandGenderifDi == 2~2,
      BiologicalSexandGenderifDi == 4~2,
      BiologicalSexandGenderifDi == 5~1,
    )
  )
df$BiologicalSex <- as.factor(df$BiologicalSex)

```

## Dealing with the medication info table

```{r}
library(readxl)
library(dplyr)

meds <- read_excel("Salience_S51TTR_TxMeds.xlsx", sheet = "MedsDuringTMS")
names(meds)[1] <- "RID"
meds_selected <- meds %>%
  inner_join(df, by = "RID")

# SSRI
meds_selected <- meds_selected %>%
  rowwise() %>%
  mutate(SSRI_use = if_else(any(!is.na(c_across(c(
    `Fluoxetine / Prozac®`,
     Trintellix, 
    `Citalopram / Celexa®`,
    `Fluvoxamine / Luvox®`,
    `Paroxetine / Paxil®`,
    `Sertraline / Zoloft®`,
    `Escitalopram / Lexapro®`,
    `Viibryd / Vilazodone`
    )))), 
    TRUE, FALSE))

# SNRI
meds_selected$Pristiq <- as.numeric(meds_selected$Pristiq)
meds_selected <- meds_selected %>%
  rowwise() %>%
  mutate(SNRI_use = if_else(any(!is.na(c_across(c(
    `Venlafaxine / Effexor`,
    `Duloxetine / Cymbalta®`, 
     Pristiq,
    `Fetzima / Levomilnacipran`
    )))), 
    TRUE, FALSE))

# SSRI or SNRI
meds_selected <- meds_selected %>%
  mutate(SSRI_SNRI_use = SSRI_use | SNRI_use)

# TCA
meds_selected <- meds_selected %>%
  rowwise() %>%
  mutate(TCA_use = if_else(any(!is.na(c_across(c(
    `Amitriptyline / Elavil / Endep`,
    `Imipramine / Tofranil®`, 
    `Desipramine/Norpramin/Pertofrane`,
    `Trimipramine/Surmontil`,
    `Clomipramine / Anafranil®`,
    `Maprotilene/Ludiomil`,
    `Doxepin / Sinequan®`,
    `Nortriptyline / Pamelor / Aventyl`,
    `Protriptyline/Vivactil`
    )))), 
    TRUE, FALSE))

# MAOI
meds_selected <- meds_selected %>%
  rowwise() %>%
  mutate(MAOI_use = if_else(any(!is.na(c_across(c(
    `Phenelzine / Nardil®`,
    `Selegiline / Eldepryl®`, 
    `Selegiline Transdermal Patch/Emsam`,
    `Tranylcypromine / Parnate®`,
    `Isocarboxazid / Marplan®`,
    )))), 
    TRUE, FALSE))

# other ADs
meds_selected <- meds_selected %>%
  rowwise() %>%
  mutate(OtherAD_use = if_else(any(!is.na(c_across(c(
    `Bupropion / Wellbutrin®`,
    `Mirtazapine / Remeron®`, 
    `Nefazodone / Serzone®`,
    `Trazodone / Desyrel®`,
    `Amoxapine / Asendin®`,
    )))), 
    TRUE, FALSE))

# all kinds of antidepressant
meds_selected <- meds_selected %>%
  mutate(AD_use = SSRI_use | SNRI_use | TCA_use | MAOI_use | OtherAD_use)

# benzos
meds_selected$`Xanax (Alprazolam)` <- as.numeric(meds_selected$`Xanax (Alprazolam)`)
meds_selected <- meds_selected %>%
  rowwise() %>%
  mutate(benzo_use = if_else(any(!is.na(c_across(c(
    `Xanax (Alprazolam)`,
    `Klonopin (Clonazepam)`,
    `Valium (Diazepam)`,
    `Ativan (Lorazepam)`,
    `Oxazepam/Serax`,
    `Librium (Chlordiazepoxide)`,
    `Dalmane (Flurazepam)`,
    `Restoril (Temazepam)`,
    Prazepam,
    `Tranxene (Clorazepate)`
    )))), 
    TRUE, FALSE))

# Anticonvulsant
tegretol <- c(rep(NA, nrow(meds_selected)))
tegretol[meds_selected$C_AdditionalMedName == "Tegretol (Carbamazepine)"] <- 20
meds_selected <- cbind(meds_selected, tegretol)
meds_selected <- meds_selected %>%
  rowwise() %>%
  mutate(anticonvulsant_use = if_else(any(!is.na(c_across(c(
    `Depakote (Divalproex, Valproate)`,
    `Lyrica (Pregabalin)`,
    `Neurontin (Gabapentin)`,
    `Keppra (Levetiracetam)`,
    `Trileptal (Oxcarbazepine)`,
    `Luminal (Phenobarbitol)`,
    Secobarbital,
    tegretol,
    `Lamictal (Lamotrigine)`
    )))), 
    TRUE, FALSE))

# Augmenting
zyprexa <- c(rep(NA, nrow(meds_selected)))
zyprexa[meds_selected$A_AdditionalMedName == "Zyprexa (Olanzapine)"] <- 20
meds_selected <- cbind(meds_selected, zyprexa)

rexulti <- c(rep(NA, nrow(meds_selected)))
rexulti[meds_selected$A_AdditionalMedName == "Rexulti (Brexpiprazole)"] <- 20
meds_selected <- cbind(meds_selected, rexulti)

risperdal <- c(rep(NA, nrow(meds_selected)))
risperdal[meds_selected$A_AdditionalMedName == "Risperdal (Risperidone)"] <- 20
meds_selected <- cbind(meds_selected, risperdal)

meds_selected <- meds_selected %>%
  rowwise() %>%
  mutate(Augmenting_use = if_else(any(!is.na(c_across(c(
    `Abilify (Aripiprazole)`,
    `Lithium / Eskalith® / Lithobid®`,
    `Latuda (Lurasidone)`,
    `Seroquel (Quetipine)`,
    `Geodon (Ziprasidone)`,
     zyprexa, rexulti, risperdal
    )))), 
    TRUE, FALSE))

meds_selected <- meds_selected %>%
  rowwise() %>%
  mutate(Lithium_use = if_else(any(!is.na(c_across(c(
    `Lithium / Eskalith® / Lithobid®`
    )))), 
    TRUE, FALSE))

# conbine medication info into main results df
df <- cbind(df, 
            meds_selected$SSRI_use, 
            meds_selected$SNRI_use, 
            meds_selected$TCA_use,
            meds_selected$MAOI_use,
            meds_selected$OtherAD_use, 
            meds_selected$benzo_use, 
            meds_selected$anticonvulsant_use, 
            meds_selected$Augmenting_use, 
            meds_selected$SSRI_SNRI_use, 
            meds_selected$AD_use,
            meds_selected$Lithium_use)

# medication changes
med_history <- meds_selected %>% select(starts_with("EndDate"))
filtered_data <- med_history %>%
  filter(if_any(everything(), ~ .x != "Current" & !is.na(.x)))
```

## Comparing clinical characteristics among 4 groups

```{r}
library(dplyr)
library(stringr)


table1 <- df %>%
  select(c(
    "AgeMTDay",
    "BiologicalSex",
    "BaselinePHQ9",
    "Anxiety",
    "X_traj_Group",
    "meds_selected$SSRI_SNRI_use",
    "meds_selected$TCA_use",
    "meds_selected$OtherAD_use",
    "meds_selected$benzo_use",
    "meds_selected$Augmenting_use",
    "meds_selected$anticonvulsant_use",
    "meds_selected$Lithium_use",
    "ofCompletedTreatments",
    "FinalTxDate",
    "MTStartDate",
    "ComorbiditiesNotDepression"
    )) 
table1$X_traj_Group <- as.factor(table1$X_traj_Group)
table1 <- table1 %>%
  mutate(Bipolar = if_else(str_detect(ComorbiditiesNotDepression, "Bipolar"), 1, 0))


# anova comparing age
result <- aov(AgeMTDay ~ X_traj_Group, data = table1)
summary(result)

# treatment lengths in weeks
FinalTxDate <- as.Date(table1$FinalTxDate, format = "%d%b%Y")
StartTxDate <- as.Date(table1$MTStartDate, format = "%d%b%Y")
table1$Txlength <- as.numeric(FinalTxDate - StartTxDate)/7
result <- aov(Txlength ~ X_traj_Group, data = table1)
summary(result)
tukey_test <- TukeyHSD(result)
print(tukey_test)

# session number
result <- aov(ofCompletedTreatments ~ X_traj_Group, data = table1)
summary(result)
tukey_test <- TukeyHSD(result)
print(tukey_test)

# baseline symptom
result <- aov(BaselinePHQ9 ~ X_traj_Group, data = table1)
summary(result)
tukey_test <- TukeyHSD(result)
print(tukey_test)

# chi squared analyses for sex ratio
chisq.test(table1$X_traj_Group, table1$BiologicalSex, correct=FALSE)

# anxiety
chisq.test(table1$X_traj_Group, table1$Anxiety, correct=FALSE)

# bipolar
chisq.test(table1$X_traj_Group, table1$Bipolar, correct=FALSE)

# medication
chisq.test(table1$X_traj_Group, table1$`meds_selected$SSRI_SNRI_use`, correct=FALSE)
chisq.test(table1$X_traj_Group, table1$`meds_selected$TCA_use`, correct=FALSE)
chisq.test(table1$X_traj_Group, table1$`meds_selected$OtherAD_use`, correct=FALSE)
chisq.test(table1$X_traj_Group, table1$`meds_selected$benzo_use`, correct=FALSE)
chisq.test(table1$X_traj_Group, table1$`meds_selected$Augmenting_use`, correct=FALSE)
chisq.test(table1$X_traj_Group, table1$`meds_selected$anticonvulsant_use`, correct=FALSE)
chisq.test(table1$X_traj_Group, table1$`meds_selected$Lithium_use`, correct=FALSE)
```

## Make Table 1

```{r}
library(gtsummary)
library(officer)

table1 <- df %>%
  select(c(
    "AgeMTDay",
    "BiologicalSex",
    "BaselinePHQ9",
    "Anxiety",
    "X_traj_Group",
    "meds_selected$SSRI_SNRI_use",
    "meds_selected$TCA_use",
    "meds_selected$OtherAD_use",
    "meds_selected$benzo_use",
    "meds_selected$Augmenting_use",
    "meds_selected$anticonvulsant_use",
    "meds_selected$Lithium_use"
    )) 

table1$BiologicalSex <- factor(
                      table1$BiologicalSex,
                      levels = c(1:2),
                      labels = c("Female", "Male")
                        )
table1$X_traj_Group <- factor(
                      table1$X_traj_Group,
                      levels = c(1:4),
                      labels = c("Rapid response, standard treatment length", "Rapid response, extended plateau", "Linear response, extended course", "Nonresponse")
                        )
table1$BaselinePHQ9 <- as.integer(table1$BaselinePHQ9)
table1$Anxiety <- factor(
                    table1$Anxiety,
                    levels = c(0,1),
                    labels = c("No anxiety comorbidity", "Any anxiety comorbidity")
)

names(table1)[1] <- "Age (years)"
names(table1)[2] <- "Sex"
names(table1)[3] <- "Baseline depressive symptoms"
names(table1)[4] <- "Anxiety comorbidity"
names(table1)[6:12] <- c("SSRI/SNRI use",
                         "TCA use",
                         "Other antidepressant use",
                         "Benzodiazepine use",
                         "Augmenting agent use",
                         "Anticonvulsant use",
                         "Lithium use"
                         )

group_summary <- table1 %>% 
  tbl_summary(by = "X_traj_Group",
            statistic = list(all_continuous() ~ "{mean} ({sd})"),
            digits = list(all_continuous() ~ c(2, 2), all_categorical() ~ c(0, 2))
            ) %>% 
  add_overall(col_label = "**Total sample**, N = {N}")

group_summary
```

## To build a supplementary table showing clinical characteristics of BD patients

```{r}
library(gtsummary)
library(officer)
library(dplyr)
library(stringr)

filtered_df <- df %>% filter(str_detect(ComorbiditiesNotDepression, "Bipolar"))

table1 <- filtered_df %>%
  select(c(
    "AgeMTDay",
    "BiologicalSex",
    "BaselinePHQ9",
    "Anxiety",
    "X_traj_Group",
    "meds_selected$SSRI_SNRI_use",
    "meds_selected$TCA_use",
    "meds_selected$OtherAD_use",
    "meds_selected$benzo_use",
    "meds_selected$Augmenting_use",
    "meds_selected$anticonvulsant_use",
    "meds_selected$Lithium_use"
    )) 

table1$BiologicalSex <- factor(
                      table1$BiologicalSex,
                      levels = c(1:2),
                      labels = c("Female", "Male")
                        )
table1$X_traj_Group <- factor(
                      table1$X_traj_Group,
                      levels = c(1:4),
                      labels = c("Rapid response, standard treatment length", "Slowed response", "Linear response, extended course", "Nonresponse")
                        )
table1$BaselinePHQ9 <- as.integer(table1$BaselinePHQ9)
table1$Anxiety <- factor(
                    table1$Anxiety,
                    levels = c(0,1),
                    labels = c("No anxiety comorbidity", "Any anxiety comorbidity")
)

names(table1)[1] <- "Age (years)"
names(table1)[2] <- "Sex"
names(table1)[3] <- "Baseline depressive symptoms"
names(table1)[4] <- "Anxiety comorbidity"
names(table1)[6:12] <- c("SSRI/SNRI use",
                         "TCA use",
                         "Other antidepressant use",
                         "Benzodiazepine use",
                         "Augmenting agent use",
                         "Anticonvulsant use",
                         "Lithium use"
                         )

group_summary <- table1 %>% 
  tbl_summary(by = "X_traj_Group",
            statistic = list(all_continuous() ~ "{mean} ({sd})"),
            digits = list(all_continuous() ~ c(2, 2), all_categorical() ~ c(0, 2))
            ) %>% 
  add_overall(col_label = "**Total sample**, N = {N}")

group_summary
```

## To build a supplementary table showing clinical characteristics of patients with \> 36 session vs. 36 sessions and less

```{r}
library(gtsummary)
library(officer)
library(dplyr)
library(stringr)

df_session36 <- df %>% mutate(group_session = ifelse(ofCompletedTreatments > 36, 1, 2))

table1 <- df_session36 %>%
  select(c(
    "AgeMTDay",
    "BiologicalSex",
    "BaselinePHQ9",
    "Anxiety",
    "group_session",
    "meds_selected$SSRI_SNRI_use",
    "meds_selected$TCA_use",
    "meds_selected$OtherAD_use",
    "meds_selected$benzo_use",
    "meds_selected$Augmenting_use",
    "meds_selected$anticonvulsant_use",
    "meds_selected$Lithium_use"
    )) 

table1$BiologicalSex <- factor(
                      table1$BiologicalSex,
                      levels = c(1:2),
                      labels = c("Female", "Male")
                        )
table1$group_session <- factor(
                      table1$group_session,
                      levels = c(1:2),
                      labels = c("Extended course", "Standard course")
                        )
table1$BaselinePHQ9 <- as.integer(table1$BaselinePHQ9)
table1$Anxiety <- factor(
                    table1$Anxiety,
                    levels = c(0,1),
                    labels = c("No anxiety comorbidity", "Any anxiety comorbidity")
)

names(table1)[1] <- "Age (years)"
names(table1)[2] <- "Sex"
names(table1)[3] <- "Baseline depressive symptoms"
names(table1)[4] <- "Anxiety comorbidity"
names(table1)[6:12] <- c("SSRI/SNRI use",
                         "TCA use",
                         "Other antidepressant use",
                         "Benzodiazepine use",
                         "Augmenting agent use",
                         "Anticonvulsant use",
                         "Lithium use"
                         )

group_summary <- table1 %>% 
  tbl_summary(by = "group_session",
            statistic = list(all_continuous() ~ "{mean} ({sd})"),
            digits = list(all_continuous() ~ c(2, 2), all_categorical() ~ c(0, 2))
            ) %>% 
  add_overall(col_label = "**Total sample**, N = {N}")

group_summary
```

## To do multinomial logistic regression

```{r}
library(nnet)
library(tidyr)
library(dplyr)

# select data
df <- df %>% 
  mutate(across(c(`meds_selected$AD_use`,
                  `meds_selected$benzo_use`,
                  `meds_selected$Augmenting_use`,
                  `meds_selected$anticonvulsant_use`,
                  `meds_selected$SSRI_SNRI_use`,
                  `meds_selected$SSRI_use`,
                  `meds_selected$SNRI_use`,
                  `meds_selected$OtherAD_use`
                  ), 
                ~ifelse(.x, 1, 0)
                )) %>% 
  mutate(W2delta = Y - W2PHQ9,
         W2relative_delta = (Y - W2PHQ9)/Y) %>% 
  mutate(W1delta = Y - W1PHQ9,
         W1relative_delta = (Y - W1PHQ9)/Y)

# to build a function to extract p values
pValue_extract <- function(x){
  z <- summary(x)$coefficients/summary(x)$standard.errors
  # 2-tailed Wald z tests to test significance of coefficients
  p <- (1 - pnorm(abs(z), 0, 1)) * 2
  p
}

df$X_traj_Group <- as.factor(df$X_traj_Group)
df$X_traj_Group <- relevel(df$X_traj_Group, ref = 2)
df$BiologicalSexandGenderifDi <- as.factor(df$BiologicalSexandGenderifDi)
df$Race <- as.factor(df$Race)
df$ofMedsDuringTx <- as.factor(df$ofMedsDuringTx)


model <- multinom(X_traj_Group ~ AgeMTDay + BaselinePHQ9 + BiologicalSex  + Anxiety, data = df, weights = df$Group_Weights)

coef_estimates <- coef(model)
std_errors <- sqrt(diag(vcov(model)))
conf_interval <- 1.96
lower_bound <- coef_estimates - (conf_interval * std_errors)
upper_bound <- coef_estimates + (conf_interval * std_errors)
odds_ratios <- exp(coef_estimates)
ci_lower <- exp(lower_bound)
ci_upper <- exp(upper_bound)
odds_ratios
ci_lower
ci_upper

model <- multinom(X_traj_Group ~ AgeMTDay +BiologicalSex  + Anxiety + W1relative_delta, data = df, weights = df$Group_Weights)

coef_estimates <- coef(model)
std_errors <- sqrt(diag(vcov(model)))
conf_interval <- 1.96
lower_bound <- coef_estimates - (conf_interval * std_errors)
upper_bound <- coef_estimates + (conf_interval * std_errors)
odds_ratios <- exp(coef_estimates)
ci_lower <- exp(lower_bound)
ci_upper <- exp(upper_bound)
odds_ratios
ci_lower
ci_upper

# # original version: flagged for problemtic way of calculating CIs of OR!
# OR_std <- exp(coef(model))*summary(model)$standard.errors
# CI_upper <- exp(coef(model)) + 1.96*OR_std
# CI_lower <- exp(coef(model)) - 1.96*OR_std
# pValue_extract(model)# p values
# exp(coef(model))# OR
# CI_lower# CI
# CI_upper

```

## calculate the c-statistics and Hosmer-Lemeshow test results

```{r}
# install and library all packages
# install.packages("pROC")
# install.packages("ResourceSelection")
# Load the package
library(pROC)
library(ResourceSelection)
library(tidyr)
library(dplyr)

results <- data.frame(matrix(nrow = 6, ncol = 3))
names(results)[2] <- "c-statisitcs"
names(results)[3] <- "H-L test's p"
temp <- 0
for (i in 1:3){
  temp2 <- i+1
  for (j in temp2:4){
    temp <- temp + 1
    table_2group <- df %>%
      filter(X_traj_Group == i | X_traj_Group == j) %>%
      mutate(X_traj_Group = case_when(
              X_traj_Group == i~0,
              X_traj_Group == j~1,
        )
      )
    
model <- glm(X_traj_Group ~ AgeMTDay + BaselinePHQ9 + BiologicalSex + Anxiety, data = table_2group, family = "binomial")
    
# Calculate predicted probabilities
predicted_probabilities <- predict(model, type = "response")
    
# Calculate the c-statistic
roc_obj <- roc(table_2group$X_traj_Group, predicted_probabilities)
    c_statistic <- auc(roc_obj)
    
    # Calculate observed probabilities
    observed_probabilities <- table_2group$X_traj_Group
    hosmer_lemeshow_test <- hoslem.test(observed_probabilities, predicted_probabilities)
    
    # save out results 
    results[temp,1] <- paste(as.character(i), as.character(j), sep = "-")
    results[temp,2] <- c_statistic
    results[temp,3] <- hosmer_lemeshow_test$p.value
  }
}
```
