---
title: "main_analysis"
output:
  html_document: default
  pdf_document: default
date: "2023-03-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Main analyses of the trajectory project Based on the four-group model

group 1,2,4 cubic, group 3 quadratic. Comparing the group info of week 7 and week 10 models 2. do stats on the number of missing values of each

```{r}
# import data
rm(list = ls())

trajectory_data_clean <- read.csv("Trajectory_Data_Clean.csv")
wk10_group <- read.csv("wk10_model_group.csv")
names(wk10_group) <- c("wk10_model_group")
wk7_group <- read.csv("wk7_model_group.csv")
names(wk7_group) <- c("wk7_model_group")

# append group variables to data.csv
trajectory_data_clean <- cbind(trajectory_data_clean, wk10_group)
trajectory_data_clean <- cbind(trajectory_data_clean, wk7_group)
```

## Getting the cumulative response and remission rates

```{r}
library(tidyverse)
library(dplyr)
library(compute.es)
library(effsize)

# For supplementary table, calculate the remission and response rates for all 4 groups
# select data
data_insted <- trajectory_data_clean %>%
                select(c(
                        "Y","W1PHQ9", "W2PHQ9", "W3PHQ9", "W4PHQ9", "W5PHQ9",
                        "W6PHQ9", "W7PHQ9", "W8PHQ9", "W9PHQ9", "W10PHQ9",
                        "wk10_model_group"
                        )
                       ) 

W1_all <- data_insted %>% 
       summarize(remission_n_all  = sum(W1PHQ9 < 5, na.rm = TRUE),
                 response_n_all = sum((Y - W1PHQ9) / Y >= 0.5, na.rm = TRUE),
                 remission_rate_all = sum(W1PHQ9 < 5, na.rm = TRUE) / n(),
                 response_rate_all = sum((Y - W1PHQ9) / Y >= 0.5, na.rm = TRUE) / n()
                 )

W2_all <- data_insted %>% 
       summarize(remission_n_all  = sum(W2PHQ9 < 5, na.rm = TRUE),
                response_n_all = sum((Y - W2PHQ9) / Y >= 0.5, na.rm = TRUE),
                remission_rate_all = sum(W2PHQ9 < 5, na.rm = TRUE) / n(),
                response_rate_all = sum((Y - W2PHQ9) / Y >= 0.5, na.rm = TRUE) / n()
                )

W3_all <- data_insted %>% 
       summarize(remission_n_all  = sum(W3PHQ9 < 5, na.rm = TRUE),
                response_n_all = sum((Y - W3PHQ9) / Y >= 0.5, na.rm = TRUE),
                remission_rate_all = sum(W3PHQ9 < 5, na.rm = TRUE) / n(),
                response_rate_all = sum((Y - W3PHQ9) / Y >= 0.5, na.rm = TRUE) / n()
                )

W4_all <- data_insted %>% 
       summarize(remission_n_all  = sum(W4PHQ9 < 5, na.rm = TRUE),
                response_n_all = sum((Y - W4PHQ9) / Y >= 0.5, na.rm = TRUE),
                remission_rate_all = sum(W4PHQ9 < 5, na.rm = TRUE) / n(),
                response_rate_all = sum((Y - W4PHQ9) / Y >= 0.5, na.rm = TRUE) / n()
                )

W5_all <- data_insted %>% 
       summarize(remission_n_all  = sum(W5PHQ9 < 5, na.rm = TRUE),
                response_n_all = sum((Y - W5PHQ9) / Y >= 0.5, na.rm = TRUE),
                remission_rate_all = sum(W5PHQ9 < 5, na.rm = TRUE) / n(),
                response_rate_all = sum((Y - W5PHQ9) / Y >= 0.5, na.rm = TRUE) / n()
                )

W6_all <- data_insted %>% 
       summarize(remission_n_all  = sum(W6PHQ9 < 5, na.rm = TRUE),
                response_n_all = sum((Y - W6PHQ9) / Y >= 0.5, na.rm = TRUE),
                remission_rate_all = sum(W6PHQ9 < 5, na.rm = TRUE) / n(),
                response_rate_all = sum((Y - W6PHQ9) / Y >= 0.5, na.rm = TRUE) / n()
                )

W7_all <- data_insted %>% 
       summarize(remission_n_all  = sum(W7PHQ9 < 5, na.rm = TRUE),
                response_n_all = sum((Y - W7PHQ9) / Y >= 0.5, na.rm = TRUE),
                remission_rate_all = sum(W7PHQ9 < 5, na.rm = TRUE) / n(),
                response_rate_all = sum((Y - W7PHQ9) / Y >= 0.5, na.rm = TRUE) / n()
                )

# For week 8-10, I simply added the response_n and remission_n in week 8-10 to the response_n and remission_n in week 7 and minus those who were reponse or remission but remain to a prolonged session
response_remain <- data_insted %>%
                  summarize(response_remain_n = sum((Y - W7PHQ9) / Y >= 0.5 & (Y - W8PHQ9) / Y >= 0.5, na.rm = TRUE)
                  )

remission_remain <- data_insted %>%
                  summarize(remission_remain_n = sum(W7PHQ9 < 5 & W8PHQ9 < 5, na.rm = TRUE)
                  )

data_insted_prolonged <- trajectory_data_clean %>%
                select(c(
                        "Y","W1PHQ9", "W2PHQ9", "W3PHQ9", "W4PHQ9", "W5PHQ9",
                        "W6PHQ9", "W7PHQ9", "W8PHQ9", "W9PHQ9", "W10PHQ9",
                        "wk10_model_group"
                        )
                       ) %>%
                filter(!is.na(W8PHQ9))

W8_all <- data_insted_prolonged %>%
                summarize(remission_n_all = sum(W8PHQ9 < 5, na.rm = TRUE),
                          response_n_all = sum((Y - W8PHQ9) / Y >= 0.5, na.rm = TRUE)
                          ) %>%
                reframe(remission_n_all = remission_n_all + W7_all$remission_n_all - remission_remain$remission_remain_n,
                        response_n_all = response_n_all + W7_all$response_n_all - response_remain$response_remain_n,
                        remission_rate_all = remission_n_all/324,
                        response_rate_all = response_n_all/324
                        )

W9_all <- data_insted_prolonged %>%
                summarize(remission_n_all = sum(W9PHQ9 < 5, na.rm = TRUE),
                          response_n_all = sum((Y - W9PHQ9) / Y >= 0.5, na.rm = TRUE)
                          ) %>%
                reframe(remission_n_all = remission_n_all + W7_all$remission_n_all - remission_remain$remission_remain_n,
                        response_n_all = response_n_all + W7_all$response_n_all - response_remain$response_remain_n,
                        remission_rate_all = remission_n_all/324,
                        response_rate_all = response_n_all/324
                        )

W10_all <- data_insted_prolonged %>%
                summarize(remission_n_all = sum(W10PHQ9 < 5, na.rm = TRUE),
                          response_n_all = sum((Y - W10PHQ9) / Y >= 0.5, na.rm = TRUE)
                          ) %>%
                reframe(remission_n_all = remission_n_all + W7_all$remission_n_all - remission_remain$remission_remain_n,
                        response_n_all = response_n_all + W7_all$response_n_all - response_remain$response_remain_n,
                        remission_rate_all = remission_n_all/324,
                        response_rate_all = response_n_all/324
                        )

# grouped results
W1 <- data_insted %>%       
  group_by(wk10_model_group) %>%
      summarize(remission_n  = sum(W1PHQ9 < 5, na.rm = TRUE),
                response_n = sum((Y - W1PHQ9) / Y >= 0.5, na.rm = TRUE),
                remission_rate = sum(W1PHQ9 < 5, na.rm = TRUE) / n(),
                response_rate = sum((Y - W1PHQ9) / Y >= 0.5, na.rm = TRUE) / n(),
                N = n()
                )

W2 <- data_insted %>% 
      group_by(wk10_model_group) %>%
      summarize(remission_n  = sum(W2PHQ9 < 5, na.rm = TRUE),
                response_n = sum((Y - W2PHQ9) / Y >= 0.5, na.rm = TRUE),
                remission_rate = sum(W2PHQ9 < 5, na.rm = TRUE) / n(),
                response_rate = sum((Y - W2PHQ9) / Y >= 0.5, na.rm = TRUE) / n())

W3 <- data_insted %>% 
      group_by(wk10_model_group) %>%
      summarize(remission_n  = sum(W3PHQ9 < 5, na.rm = TRUE),
                response_n = sum((Y - W3PHQ9) / Y >= 0.5, na.rm = TRUE),
                remission_rate = sum(W3PHQ9 < 5, na.rm = TRUE) / n(),
                response_rate = sum((Y - W3PHQ9) / Y >= 0.5, na.rm = TRUE) / n())

W4 <- data_insted %>% 
      group_by(wk10_model_group) %>%
      summarize(remission_n  = sum(W4PHQ9 < 5, na.rm = TRUE),
                response_n = sum((Y - W4PHQ9) / Y >= 0.5, na.rm = TRUE),
                remission_rate = sum(W4PHQ9 < 5, na.rm = TRUE) / n(),
                response_rate = sum((Y - W4PHQ9) / Y >= 0.5, na.rm = TRUE) / n()
                )

W5 <- data_insted %>% 
      group_by(wk10_model_group) %>%
      summarize(remission_n  = sum(W5PHQ9 < 5, na.rm = TRUE),
                response_n = sum((Y - W5PHQ9) / Y >= 0.5, na.rm = TRUE),
                remission_rate = sum(W5PHQ9 < 5, na.rm = TRUE) / n(),
                response_rate = sum((Y - W5PHQ9) / Y >= 0.5, na.rm = TRUE) / n()
                )

W6 <- data_insted %>% 
      group_by(wk10_model_group) %>%
      summarize(remission_n  = sum(W6PHQ9 < 5, na.rm = TRUE),
                response_n = sum((Y - W6PHQ9) / Y >= 0.5, na.rm = TRUE),
                remission_rate = sum(W6PHQ9 < 5, na.rm = TRUE) / n(),
                response_rate = sum((Y - W6PHQ9) / Y >= 0.5, na.rm = TRUE) / n()
                )

W7 <- data_insted %>% 
      group_by(wk10_model_group) %>%
      summarize(remission_n  = sum(W7PHQ9 < 5, na.rm = TRUE),
                response_n = sum((Y - W7PHQ9) / Y >= 0.5, na.rm = TRUE),
                remission_rate = sum(W7PHQ9 < 5, na.rm = TRUE) / n(),
                response_rate = sum((Y - W7PHQ9) / Y >= 0.5, na.rm = TRUE) / n()
                )

# get the remaining subjects who repspond to treatment or were already in remission
response_remain_group <- data_insted %>%
                  group_by(wk10_model_group) %>%
                  summarize(response_remain_n = sum((Y - W7PHQ9) / Y >= 0.5 & (Y - W8PHQ9) / Y >= 0.5, na.rm = TRUE)
                  )
remission_remain_group <- data_insted %>%
                  group_by(wk10_model_group) %>%
                  summarize(remission_remain_n = sum(W7PHQ9 < 5 & W8PHQ9 < 5, na.rm = TRUE)
                  )

W8 <- data_insted_prolonged %>%
                group_by(wk10_model_group) %>%
                summarize(remission_n = sum(W8PHQ9 < 5, na.rm = TRUE),
                          response_n = sum((Y - W8PHQ9) / Y >= 0.5, na.rm = TRUE)
                        ) 
W8 <- rbind(list(1,0,0),W8)
W8 <- W8 %>%
          reframe(wk10_model_group = wk10_model_group,
            remission_n = remission_n + W7$remission_n - remission_remain_group$remission_remain_n,
                  response_n = response_n + W7$response_n - response_remain_group$response_remain_n,
                  remission_rate = remission_n/W1$N,
                  respoinse_rate = response_n/W1$N
                        )

W9 <- data_insted_prolonged %>%
                group_by(wk10_model_group) %>%
                summarize(remission_n = sum(W9PHQ9 < 5, na.rm = TRUE),
                          response_n = sum((Y - W9PHQ9) / Y >= 0.5, na.rm = TRUE)
                        ) 
W9 <- rbind(list(1,0,0),W9)
W9 <- W9 %>%
          reframe(wk10_model_group = wk10_model_group,
            remission_n = remission_n + W7$remission_n - remission_remain_group$remission_remain_n,
                  response_n = response_n + W7$response_n - response_remain_group$response_remain_n,
                  remission_rate = remission_n/W1$N,
                  respoinse_rate = response_n/W1$N
                        )

W10 <- data_insted_prolonged %>%
                group_by(wk10_model_group) %>%
                summarize(remission_n = sum(W10PHQ9 < 5, na.rm = TRUE),
                          response_n = sum((Y - W10PHQ9) / Y >= 0.5, na.rm = TRUE)
                        ) 
W10 <- rbind(list(1,0,0),W10)
W10 <- W10 %>%
          reframe(wk10_model_group = wk10_model_group,
            remission_n = remission_n + W7$remission_n - remission_remain_group$remission_remain_n,
                  response_n = response_n + W7$response_n - response_remain_group$response_remain_n,
                  remission_rate = remission_n/W1$N,
                  respoinse_rate = response_n/W1$N
                        )

# do chi square analysis on remission and response data for all 4 groups
data_frames <- list(W1, W2, W3, W4, W5, W6, W7, W8, W9, W10)
stats_table_response = cbind(rep(0,10),rep(0,10))
stats_table_remission = cbind(rep(0,10),rep(0,10))
for (i in 1:10){
  current_table <- cbind(data_frames[[i]][,3], W1$N-data_frames[[i]][,3])
  results_chi <- chisq.test(current_table)
  stats_table_response[i,1] <- results_chi$statistic
  stats_table_response[i,2] <- results_chi$p.value
  
  current_table <- cbind(data_frames[[i]][,2], W1$N-data_frames[[i]][,2])
  results_chi <- chisq.test(current_table)
  stats_table_remission[i,1] <- results_chi$statistic
  stats_table_remission[i,2] <- results_chi$p.value
}

```

## To do secondary analysis

comparing the PHQ-9 scores, remission rates (PHQ-9 \< 5), response rate (reduction rates of PHQ-9 score ≥ 50%) at week 1, week7, and week 10 for all four groups

```{r}
library(tidyverse)
library(dplyr)
library(compute.es)
library(effsize)

# select data
data_insted <- trajectory_data_clean %>%
                select(c(
                        "Y","W1PHQ9", "W2PHQ9", "W3PHQ9", "W4PHQ9", "W5PHQ9",
                        "W6PHQ9", "W7PHQ9", "W8PHQ9", "W9PHQ9", "W10PHQ9",
                        "wk10_model_group"
                        )
                       ) %>%
                filter(!is.na(W7PHQ9) & !is.na(W10PHQ9))
# calculate the response rate and the remission rate
W7 <- data_insted %>% 
      group_by(wk10_model_group) %>%
      summarize(N = n(),
                remission_n  = sum(W7PHQ9 < 5),
                response_n = sum((Y - W7PHQ9) / Y >= 0.5),
                mean_value = mean(W7PHQ9),
                sd_value = sd(W7PHQ9),
                response_rate = sum((Y - W7PHQ9) / Y >= 0.5) / n(),
                remission_rate = sum(W7PHQ9 < 5) / n())
W10<- data_insted %>% 
      group_by(wk10_model_group) %>%
      summarize(N = n(),
                remission_n = sum(W10PHQ9 < 5),
                response_n = sum((Y - W10PHQ9) / Y >= 0.5),
                mean_value = mean(W10PHQ9),
                sd_value = sd(W10PHQ9),
                response_rate = sum((Y - W10PHQ9) / Y >= 0.5) / n(),
                remission_rate = sum(W10PHQ9 < 5) / n())

# buid a table "results" to store results
results_wk7 <- rbind(t(W7[1,5:8]), t(W7[2,5:8]), t(W7[3,5:8]))
results_wk10 <- rbind(t(W10[1,5:8]), t(W10[2,5:8]), t(W10[3,5:8]))
results <- cbind(results_wk7,results_wk10)
results <- cbind(results, c(rep(0, 12)), c(rep(0, 12)), c(rep(0, 12)))
colnames(results)[1:5] <- c('week 7', 'week 10', 't/chi2', 'p', 'cohend')

# McNemar test
remission_table_McNemar <- data_insted %>% 
      group_by(wk10_model_group) %>%
      summarize(a  = sum(W7PHQ9 < 5 & W10PHQ9 < 5),
                b = sum(W7PHQ9 < 5 & W10PHQ9 >= 5),
                c = sum(W7PHQ9 >= 5 & W10PHQ9 < 5),
                d = sum(W7PHQ9 >= 5 & W10PHQ9 >= 5)
                  )
response_table_McNemar <- data_insted %>% 
      group_by(wk10_model_group) %>%
      summarize(a  = sum((Y - W7PHQ9) / Y >= 0.5 & (Y - W10PHQ9) / Y >= 0.5),
                b = sum((Y - W7PHQ9) / Y >= 0.5 & (Y - W10PHQ9) / Y < 0.5),
                c = sum((Y - W7PHQ9) / Y < 0.5 & (Y - W10PHQ9) / Y >= 0.5),
                d = sum((Y - W7PHQ9) / Y < 0.5 & (Y - W10PHQ9) / Y < 0.5)
                  )

# group 2
data <- matrix(as.numeric(c(response_table_McNemar[1,2], 
                            response_table_McNemar[1,3],
                            response_table_McNemar[1,4],
                            response_table_McNemar[1,5])),
                            nrow = 2, ncol = 2, byrow = TRUE)
stats_chi_resp_group2 <- mcnemar.test(data)
print(stats_chi_resp_group2)

data <- matrix(as.numeric(c(remission_table_McNemar[1,2],
                            remission_table_McNemar[1,3],
                            remission_table_McNemar[1,4],
                            remission_table_McNemar[1,5])),
                            nrow = 2, ncol = 2, byrow = TRUE)
stats_chi_rem_group2 <- mcnemar.test(data)
print(stats_chi_rem_group2)

results[3,3] <- stats_chi_resp_group2$statistic
results[3,4] <- stats_chi_resp_group2$p.value
results[4,3] <- stats_chi_rem_group2$statistic
results[4,4] <- stats_chi_rem_group2$p.value

# group 3
data <- matrix(as.numeric(c(response_table_McNemar[2,2], 
                            response_table_McNemar[2,3],
                            response_table_McNemar[2,4],
                            response_table_McNemar[2,5])),
                            nrow = 2, ncol = 2, byrow = TRUE)
stats_chi_resp_group3 <- mcnemar.test(data)
print(stats_chi_resp_group3)

data <- matrix(as.numeric(c(remission_table_McNemar[2,2],
                            remission_table_McNemar[2,3],
                            remission_table_McNemar[2,4],
                            remission_table_McNemar[2,5])),
                            nrow = 2, ncol = 2, byrow = TRUE)
stats_chi_rem_group3 <- mcnemar.test(data)
print(stats_chi_rem_group3)

results[7,3] <- stats_chi_resp_group3$statistic
results[7,4] <- stats_chi_resp_group3$p.value
results[8,3] <- stats_chi_rem_group3$statistic
results[8,4] <- stats_chi_rem_group3$p.value

# do paired t-tests, group 2
data_insted_group <- data_insted %>% 
                    select(c("W7PHQ9","W10PHQ9")) %>% 
                    filter(data_insted$wk10_model_group == 2)
            
before <- na.omit(data_insted_group$W7PHQ9)
after <- na.omit(data_insted_group$W10PHQ9)
stats_t <- t.test(before, after, paired = TRUE)
results[1, 3] <- stats_t$statistic
results[1, 4] <- stats_t$p.value
cohens_d <- cohen.d(before, after, paired = TRUE)
results[1, 5] <- cohens_d$estimate

# do paired t-tests, group 3
data_insted_group <- data_insted %>% 
                    select(c("W7PHQ9","W10PHQ9")) %>% 
                    filter(data_insted$wk10_model_group == 3)
before <- na.omit(data_insted_group$W7PHQ9)
after <- na.omit(data_insted_group$W10PHQ9)
stats_t <- t.test(before, after, paired = TRUE)
results[5, 3] <- stats_t$statistic
results[5, 4] <- stats_t$p.value
cohens_d <- cohen.d(before, after, paired = TRUE)
results[5, 5] <- cohens_d$estimate
```

## Comparing Wk7 and Wk10 models

To do tabular analyses on wk7 and wk10 models, also calculating the kappa value

To perform analyses comparing the age, sex ratio, and baseline depressive symptoms between wk7 group and wk10 groups

```{r}
library(irr)

# Calculating the kappa value
trajectory_data_clean$wk10_model_groupF <- factor(trajectory_data_clean$wk10_model_group, levels = c(1,2,3,4))
trajectory_data_clean$wk7_model_groupF <- factor(trajectory_data_clean$wk7_model_group, levels = c(1,2,3,4))

observed <- table(trajectory_data_clean$wk10_model_groupF, 
      trajectory_data_clean$wk7_model_groupF)
observed

kappa2(cbind(trajectory_data_clean$wk10_model_groupF, trajectory_data_clean$wk7_model_groupF))

# Comparing the age, sex ratio, and the baseline depressive symptoms
# ages
group_label <- 1
age_wk7 <- trajectory_data_clean$AgeMTDay[wk7_group == group_label]
df_age_wk7 <- data.frame(age = age_wk7, group = rep(1, length(age_wk7)))

age_wk10 <- trajectory_data_clean$AgeMTDay[wk10_group == group_label]
df_age_wk10 <- data.frame(age = age_wk10, group = rep(2, length(age_wk10)))

df_age = rbind(df_age_wk7, df_age_wk10)
result <- t.test(age ~ group, data = df_age, var.equal = TRUE)
print(result)

# baseline depressive symptoms
dep_wk7 <- trajectory_data_clean$BaselinePHQ9[wk7_group == group_label]
df_dep_wk7 <- data.frame(dep = dep_wk7, group = rep(1, length(dep_wk7)))

dep_wk10 <- trajectory_data_clean$BaselinePHQ9[wk10_group == group_label]
df_dep_wk10 <- data.frame(dep = dep_wk10, group = rep(2, length(dep_wk10)))

df_dep = rbind(df_dep_wk7, df_dep_wk10)
result <- t.test(dep ~ group, data = df_dep, var.equal = TRUE)
print(result)

# sex ratio
df4chi <- data.frame(wk7=c(0,0), wk10=c(0,0))
a1 <- trajectory_data_clean %>%
  filter(wk7_model_group == group_label, BiologicalSexandGenderifDi == 1) %>%
  summarise(count = n())
a2 <- trajectory_data_clean %>%
  filter(wk7_model_group == group_label, BiologicalSexandGenderifDi == 2) %>%
  summarise(count = n())

b1 <- trajectory_data_clean %>%
  filter(wk10_model_group == group_label, BiologicalSexandGenderifDi == 1) %>%
  summarise(count = n())
b2 <- trajectory_data_clean %>%
  filter(wk10_model_group == group_label, BiologicalSexandGenderifDi == 2) %>%
  summarise(count = n())
contingency_table <- as.table(rbind(c(a1$count, a2$count), c(b1$count, b2$count)))
results_chi <- chisq.test(contingency_table)
print(results_chi)
```

## count the missing values of each weeks' PHQ scores

```{r}
library(tidyverse)
library(dplyr)

# Week 10 model
# count the absolute number of the missing value in each group
missing_counts_wk10 <- trajectory_data_clean %>%
  group_by(wk10_model_groupF) %>%
  select(25:35) %>%
  summarise_all(funs(sum(is.na(.))))

# get the number of participants in each group
group_num_wk10 <- count(trajectory_data_clean,wk10_model_groupF)
group_num_wk10

# do chi square test at all different time points
stats_table <- cbind(rep(0, 10), rep(0, 10), rep(0, 10), rep(0, 10), rep(0, 10), rep(0, 10))
colnames(stats_table) <- c('n_missing1', 'n_missing2', 'n_missing3', 'n_missing4','stats','p')
for (iTime in 3:12){
  current_missing_counts <- missing_counts_wk10[,iTime]
  current_nonMissing_counts <- group_num_wk10[,2]-current_missing_counts
  current_table <- cbind(current_missing_counts, current_nonMissing_counts)
  results_chi <- chisq.test(current_table)
  stats_table[iTime-2,5] <- results_chi$statistic
  stats_table[iTime-2,6] <- results_chi$p.value
  stats_table[iTime-2,1:4] <- t(current_missing_counts)
}

# get the relative missing value number to the absolute number of subjects
missing_counts_wk10z <- missing_counts_wk10 %>%
  select(2:ncol(.)) %>%
  mutate_all(~./group_num_wk10$n)

# add a group and model label
missing_counts_wk10z <- cbind(group_num_wk10$wk10_model_groupF,missing_counts_wk10z)
names(missing_counts_wk10z)[1] <- "Group"

# coverting wide data to long data
missing_counts_wk10z_long <- missing_counts_wk10z %>%
  pivot_longer(
    cols = 2:ncol(.),
    names_to = "Weeks",
    values_to = "Missing_z"
  )
missing_counts_wk10z_long$Time <- factor(
          missing_counts_wk10z_long$Weeks, 
          levels = c("Y","W1PHQ9","W2PHQ9","W3PHQ9","W4PHQ9","W5PHQ9","W6PHQ9","W7PHQ9","W8PHQ9","W9PHQ9","W10PHQ9"), 
          labels = as.character(c(0:10))
          )

missing_counts_wk10z_long$Group <- factor(
                        missing_counts_wk10z_long$Group, 
                        levels = c(1,2,3,4),
                        labels = c("Rapid response, standard treatment length", "Slowed response", "Linear response, extended course", "Nonresponse"))

# Week 7 model
# count the absolute number of the missing value in each group
missing_counts_wk7 <- trajectory_data_clean %>%
  group_by(wk7_model_groupF) %>%
  select(25:35) %>%
  summarise_all(funs(sum(is.na(.))))

# get the number of participants in each group
group_num_wk7 <- count(trajectory_data_clean,wk7_model_groupF)
group_num_wk7

# get the relative missing value number to the absolute number of subjects
missing_counts_wk7z <- missing_counts_wk7 %>%
  select(2:ncol(.)) %>%
  mutate_all(~./group_num_wk7$n)

# add a group label
missing_counts_wk7z <- cbind(group_num_wk7$wk7_model_groupF,missing_counts_wk7z)
names(missing_counts_wk7z)[1] <- "Group"

# coverting wide data to long data
missing_counts_wk7z_long <- missing_counts_wk7z %>%
  pivot_longer(
    cols = 2:ncol(.),
    names_to = "Weeks",
    values_to = "Missing_z"
  )
missing_counts_wk7z_long$Time <- factor(
                  missing_counts_wk7z_long$Weeks, 
                  levels = c("Y","W1PHQ9","W2PHQ9","W3PHQ9","W4PHQ9","W5PHQ9","W6PHQ9","W7PHQ9","W8PHQ9","W9PHQ9","W10PHQ9"), 
                  labels = c(as.character(c(0:10)))
                  )
missing_counts_wk7z_long$Group <- factor(
                    missing_counts_wk7z_long$Group, 
                    levels = c(1,2,3,4),
                    labels = c("Rapid response", "Response, plateau", "Linear response", "Nonresponse")
                    )
```

## To plot some supplementary figures

the bar plots and the line plot showing the ratios of missing values in each group

```{r}
# plot with ggplot2
library(ggplot2)
library(patchwork)

# plotting a bar plot showing each group's subject number
# week 10 model
ggplot(group_num_wk10, aes(x = wk10_model_groupF, y = n)) +
  geom_bar(stat = "identity") +
  xlab("Group") +
  ylab("Number of subjects") +
  ggtitle("Week 10 model") +
  theme_classic(base_size = 20)
ggsave("wk10model_subject_n.png", width = 10, height = 10, dpi = 600)

#week 7 model
ggplot(group_num_wk7, aes(x = wk7_model_groupF, y = n)) +
  geom_bar(stat = "identity") +
  xlab("Group") +
  ylab("Number of subjects") +
  ggtitle("Week 7 model") +
  theme_classic(base_size = 20)
ggsave("wk7model_subject_n.png", width = 10, height = 10, dpi = 600)

# plotting the line plots showing the missing percentage of each groups
# week 10 model
ggplot(missing_counts_wk10z_long, aes(x = Time, y = Missing_z, group = Group, color = Group)) +
  geom_line(size = 1.5) +
  labs(x = "Week", y = "Ratios of missing values", color = "Group") +
  theme_classic(base_size = 20) +
  scale_color_manual(values=c('#FF0060','#0079FF','#00DFA2','#FFE194'))+
  theme(legend.text = element_text(size = 16))

ggsave("ratio_missing_value.png", width = 13, height = 8, dpi = 600)

# week 7 model 

# p2 <- ggplot(missing_counts_wk7z_long, aes(x = Time, y = Missing_z, group = Group, color = Group)) + # geom_line(size = 1) + # labs(x = "Time", y = "Relative number of Missing Values", color = "Group") + # ggtitle("Relative number of missing values (Week 7 data)") + # theme_classic(base_size = 20) + # theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) # # ggsave("wk10model_relative_missing_value.png", width = 10, height = 8, dpi = 600)

```

## Plotting dot and line figures with shaded error bars, main results

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)

#+++++++++++++++++++++++++
# Function to calculate the mean and the standard deviation
  # for each group
#+++++++++++++++++++++++++
# data : a data frame
# varname : the name of a column containing the variable
  #to be summariezed
# groupnames : vector of column names to be used as
  # grouping variables
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}

# select data
data_dot_line <- trajectory_data_clean %>%
                select(c(
                        "Y","W1PHQ9", "W2PHQ9", "W3PHQ9", "W4PHQ9", "W5PHQ9",
                        "W6PHQ9", "W7PHQ9", "W8PHQ9", "W9PHQ9", "W10PHQ9",
                        "wk10_model_group"
                        )
                       )
# exclude group 4 from week 8-10
data_dot_line <- data_dot_line %>% 
                mutate(W8PHQ9 = ifelse(wk10_model_group == 4, NA, W8PHQ9),
                       W9PHQ9 = ifelse(wk10_model_group == 4, NA, W8PHQ9),
                       W10PHQ9 = ifelse(wk10_model_group == 4, NA, W10PHQ9)
                       )
names(data_dot_line)[1:11] <- as.character(c(0:10))
data_dtl_long <- data_dot_line %>%
              pivot_longer(
                cols = as.character(c(0:10)),
                names_to = "Time",
                values_to = "Value"
              )
df_summary <- data_summary(data_dtl_long, varname="Value", 
                    groupnames=c("Time", "wk10_model_group"))
names(df_summary)[2] <- "Group"
df_summary$Group <- as.factor(df_summary$Group)
df_summary$Group <- factor(df_summary$Group, levels = c("1", "2", "3", "4"), labels = c("Rapid response, standard treatment length", "Slowed response", "Linear response, extended course", "Nonresponse"))

# dot/line plot of week 10 model
ggplot(df_summary, aes(x= as.integer(Time), y=Value, group=Group, color=Group)) + 
  geom_line(size = 1) + xlab("Week") + ylab("PHQ-9 score") +
  geom_point(size = 3) +
  geom_ribbon(aes(ymin=Value-sd, ymax=Value+sd), fill = "grey70", alpha = .4, colour = NA) +
  scale_y_continuous(limits = c(-0.5, 30), breaks = seq(0, 30, by = 5)) +
  scale_x_continuous(limits = c(-0.5, 11), breaks = c(0:10)) +
  scale_color_manual(values=c('#FF0060','#0079FF','#00DFA2','#FFE194')) +
  theme_classic(base_size = 20) + 
  theme(legend.text = element_text(size = 12))

ggsave("wk10model_trajectory.tiff", width = 10, height = 5, dpi = 600)


# select data
data_dot_line <- trajectory_data_clean %>%
                select(c(
                        "Y","W1PHQ9", "W2PHQ9", "W3PHQ9", "W4PHQ9", "W5PHQ9",
                        "W6PHQ9", "W7PHQ9",
                        "wk7_model_group"
                        )
                       )
names(data_dot_line)[1:8] <- as.character(c(0:7))
data_dtl_long <- data_dot_line %>%
              pivot_longer(
                cols = as.character(c(0:7)),
                names_to = "Time",
                values_to = "Value"
              )
df_summary <- data_summary(data_dtl_long, varname="Value", 
                    groupnames=c("Time", "wk7_model_group"))
names(df_summary)[2] <- "Group"
df_summary$Group <- as.factor(df_summary$Group)
df_summary$Group <- factor(df_summary$Group, levels = c("1", "2", "3", "4"), labels = c("Rapid response, standard treatment length", "Slowed response", "Linear response, extended course", "Nonresponse"))

# dot/line plot of week 7 model
ggplot(df_summary, aes(x= as.integer(Time), y=Value, group=Group, color=Group)) + 
  geom_line(size = 1) + xlab("Week") + ylab("PHQ-9 score") +
  geom_point(size = 3) +
  geom_ribbon(aes(ymin=Value-sd, ymax=Value+sd), fill = "grey70", alpha = .4, colour = NA) +
  scale_y_continuous(limits = c(-0.5, 30), breaks = seq(0, 30, by = 5)) +
  scale_x_continuous(limits = c(-0.5, 8), breaks = c(0:7)) +
  scale_color_manual(values=c('#FF0060','#0079FF','#00DFA2','#FFE194')) +
  theme_classic(base_size = 20) + 
  theme(legend.text = element_text(size = 12))

ggsave("wk7model_trajectory.png", width = 10, height = 5, dpi = 600)
```
