---
title: "main_analysis"
output:
  html_document: default
  pdf_document: default
date: "2023-03-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Main analyses of the trajectory project Based on the four-group model

group 1,2,4 cubic, group 3 quadratic. Comparing the group info of week 7 and week 10 models 2. do stats on the number of missing values of each

```{r}
# import data
rm(list = ls())

trajectory_data_clean <- read.csv("Trajectory_Data_Clean.csv")
wk10_group <- read.csv("wk10_model_group.csv")
names(wk10_group) <- c("wk10_model_group")
wk7_group <- read.csv("wk7_model_group.csv")
names(wk7_group) <- c("wk7_model_group")

# append group variables to data.csv
trajectory_data_clean <- cbind(trajectory_data_clean, wk10_group)
trajectory_data_clean <- cbind(trajectory_data_clean, wk7_group)
```

## do secondary analysis

comparing the PHQ-9 scores, remission rates (PHQ-9 \< 5), response rate (reduction rates of PHQ-9 score ≥ 50%) at week 1, week7, and week 10 for all four groups

```{r}
library(tidyverse)
library(dplyr)
# install.packages("compute.es")  # Run this line if you haven't installed the package
library(compute.es)
library(effsize)

# select data
data_insted <- trajectory_data_clean %>%
                select(c(
                        "Y","W1PHQ9", "W2PHQ9", "W3PHQ9", "W4PHQ9", "W5PHQ9",
                        "W6PHQ9", "W7PHQ9", "W8PHQ9", "W9PHQ9", "W10PHQ9",
                        "wk10_model_group"
                        )
                       ) %>%
                filter(!is.na(W1PHQ9) & !is.na(W8PHQ9) & !is.na(W10PHQ9))

# calculate the response rate and the remission rate
W1 <- data_insted %>% 
      group_by(wk10_model_group) %>%
      summarize(N = n(),
                remission_n = sum(W1PHQ9 < 5),
                response_n = sum((Y - W1PHQ9) / Y >= 0.5),
                mean_value = mean(W1PHQ9),
                sd_value = sd(W1PHQ9),
                response_rate = sum(W1PHQ9 < 5) / n(),
                remission_rate = sum((Y - W1PHQ9) / Y >= 0.5) / n())

W7 <- data_insted %>% 
      group_by(wk10_model_group) %>%
      summarize(N = n(),
                remission_n  = sum(W7PHQ9 < 5),
                response_n = sum((Y - W7PHQ9) / Y >= 0.5),
                mean_value = mean(W7PHQ9),
                sd_value = sd(W7PHQ9),
                response_rate = sum((Y - W7PHQ9) / Y >= 0.5) / n(),
                remission_rate = sum(W7PHQ9 < 5) / n())
W10<- data_insted %>% 
      group_by(wk10_model_group) %>%
      summarize(N = n(),
                remission_n = sum(W10PHQ9 < 5),
                response_n = sum((Y - W10PHQ9) / Y >= 0.5),
                mean_value = mean(W10PHQ9),
                sd_value = sd(W10PHQ9),
                response_rate = sum((Y - W10PHQ9) / Y >= 0.5) / n(),
                remission_rate = sum(W10PHQ9 < 5) / n())

# buid a table "results" to store results
results_wk7 <- rbind(t(W7[1,5:8]), t(W7[2,5:8]), t(W7[3,5:8]))
results_wk10 <- rbind(t(W10[1,5:8]), t(W10[2,5:8]), t(W10[3,5:8]))
results <- cbind(results_wk7,results_wk10)
results <- cbind(results, c(rep(0, 12)), c(rep(0, 12)), c(rep(0, 12)))
colnames(results)[1:5] <- c('week 7', 'week 10', 't/chi2', 'p', 'cohend')

# # chi-square test to compare group differences, might be problematic!!
# response_table <- cbind(W10$response_n[1:2], W10$N[1:2] - W10$response_n[1:2])
# result <- chisq.test(response_table)
# 
# remission_table <- cbind(W10$remission_n[1:2], W10$N[1:2] - W10$remission_n[1:2])
# result <- chisq.test(remission_table)
# print(result)
# 

# McNemar test
remission_table_McNemar <- data_insted %>% 
      group_by(wk10_model_group) %>%
      summarize(a  = sum(W7PHQ9 < 5 & W10PHQ9 < 5),
                b = sum(W7PHQ9 < 5 & W10PHQ9 >= 5),
                c = sum(W7PHQ9 >= 5 & W10PHQ9 < 5),
                d = sum(W7PHQ9 >= 5 & W10PHQ9 >= 5)
                  )
response_table_McNemar <- data_insted %>% 
      group_by(wk10_model_group) %>%
      summarize(a  = sum((Y - W7PHQ9) / Y >= 0.5 & (Y - W10PHQ9) / Y >= 0.5),
                b = sum((Y - W7PHQ9) / Y >= 0.5 & (Y - W10PHQ9) / Y < 0.5),
                c = sum((Y - W7PHQ9) / Y < 0.5 & (Y - W10PHQ9) / Y >= 0.5),
                d = sum((Y - W7PHQ9) / Y < 0.5 & (Y - W10PHQ9) / Y < 0.5)
                  )
# # group 1
# data <- matrix(as.numeric(c(response_table_McNemar[1,2],
#                             response_table_McNemar[1,3],
#                             response_table_McNemar[1,4], 
#                             response_table_McNemar[1,5])),
#                             nrow = 2, ncol = 2, byrow = TRUE)
# stats_chi_resp_group1 <- mcnemar.test(data)
# print(stats_chi_resp_group1)
# 
# data <- matrix(as.numeric(c(remission_table_McNemar[1,2], 
#                             remission_table_McNemar[1,3],
#                             remission_table_McNemar[1,4], 
#                             remission_table_McNemar[1,5])),
#                             nrow = 2, ncol = 2, byrow = TRUE)
# stats_chi_rem_group1 <- mcnemar.test(data)
# print(stats_chi_rem_group1)
# 
# results[3,3] <- stats_chi_resp_group1$statistic
# results[3,4] <- stats_chi_resp_group1$p.value
# results[4,3] <- stats_chi_rem_group1$statistic
# results[4,4] <- stats_chi_rem_group1$p.value

# group 2
data <- matrix(as.numeric(c(response_table_McNemar[1,2], 
                            response_table_McNemar[1,3],
                            response_table_McNemar[1,4],
                            response_table_McNemar[1,5])),
                            nrow = 2, ncol = 2, byrow = TRUE)
stats_chi_resp_group2 <- mcnemar.test(data)
print(stats_chi_resp_group2)

data <- matrix(as.numeric(c(remission_table_McNemar[1,2],
                            remission_table_McNemar[1,3],
                            remission_table_McNemar[1,4],
                            remission_table_McNemar[1,5])),
                            nrow = 2, ncol = 2, byrow = TRUE)
stats_chi_rem_group2 <- mcnemar.test(data)
print(stats_chi_rem_group2)

results[3,3] <- stats_chi_resp_group2$statistic
results[3,4] <- stats_chi_resp_group2$p.value
results[4,3] <- stats_chi_rem_group2$statistic
results[4,4] <- stats_chi_rem_group2$p.value

# group 3
data <- matrix(as.numeric(c(response_table_McNemar[2,2], 
                            response_table_McNemar[2,3],
                            response_table_McNemar[2,4],
                            response_table_McNemar[2,5])),
                            nrow = 2, ncol = 2, byrow = TRUE)
stats_chi_resp_group3 <- mcnemar.test(data)
print(stats_chi_resp_group3)

data <- matrix(as.numeric(c(remission_table_McNemar[2,2],
                            remission_table_McNemar[2,3],
                            remission_table_McNemar[2,4],
                            remission_table_McNemar[2,5])),
                            nrow = 2, ncol = 2, byrow = TRUE)
stats_chi_rem_group3 <- mcnemar.test(data)
print(stats_chi_rem_group3)

results[7,3] <- stats_chi_resp_group3$statistic
results[7,4] <- stats_chi_resp_group3$p.value
results[8,3] <- stats_chi_rem_group3$statistic
results[8,4] <- stats_chi_rem_group3$p.value

# # do paired t-tests, group 1
# data_insted_group <- data_insted %>% 
#                     select(c("W7PHQ9","W10PHQ9")) %>% 
#                     filter(data_insted$wk10_model_group == 1)
#             
# before <- na.omit(data_insted_group$W7PHQ9)
# after <- na.omit(data_insted_group$W10PHQ9)
# stats_t <- t.test(before, after, paired = TRUE)
# results[1, 3] <- stats_t$statistic
# results[1, 4] <- stats_t$p.value
# cohens_d <- cohen.d(before, after, paired = TRUE)
# results[1, 5] <- cohens_d$estimate

# do paired t-tests, group 2
data_insted_group <- data_insted %>% 
                    select(c("W7PHQ9","W10PHQ9")) %>% 
                    filter(data_insted$wk10_model_group == 2)
            
before <- na.omit(data_insted_group$W7PHQ9)
after <- na.omit(data_insted_group$W10PHQ9)
stats_t <- t.test(before, after, paired = TRUE)
results[1, 3] <- stats_t$statistic
results[1, 4] <- stats_t$p.value
cohens_d <- cohen.d(before, after, paired = TRUE)
results[1, 5] <- cohens_d$estimate

# do paired t-tests, group 3
data_insted_group <- data_insted %>% 
                    select(c("W7PHQ9","W10PHQ9")) %>% 
                    filter(data_insted$wk10_model_group == 3)
before <- na.omit(data_insted_group$W7PHQ9)
after <- na.omit(data_insted_group$W10PHQ9)
stats_t <- t.test(before, after, paired = TRUE)
results[5, 3] <- stats_t$statistic
results[5, 4] <- stats_t$p.value
cohens_d <- cohen.d(before, after, paired = TRUE)
results[5, 5] <- cohens_d$estimate
```

## do tabular analyses on wk7 and wk10 models, also calculating the kappa value

```{r}
# install.packages("irr")
library(irr)

trajectory_data_clean$wk10_model_groupF <- factor(trajectory_data_clean$wk10_model_group, levels = c(1,2,3,4))
trajectory_data_clean$wk7_model_groupF <- factor(trajectory_data_clean$wk7_model_group, levels = c(1,2,3,4))

observed <- table(trajectory_data_clean$wk10_model_groupF, 
      trajectory_data_clean$wk7_model_groupF)
observed

kappa2(cbind(trajectory_data_clean$wk10_model_groupF, trajectory_data_clean$wk7_model_groupF))
```

## count the missing values of each weeks' PHQ scores

```{r}
# install.packages("tidyverse")
library(tidyverse)
library(dplyr)

# Week 10 model
# count the absolute number of the missing value in each group
missing_counts_wk10 <- trajectory_data_clean %>%
  group_by(wk10_model_groupF) %>%
  select(25:35) %>%
  summarise_all(funs(sum(is.na(.))))

# get the number of participants in each group
group_num_wk10 <- count(trajectory_data_clean,wk10_model_groupF)
group_num_wk10

# do chi square test at all different time points
stats_table <- cbind(rep(0, 10), rep(0, 10), rep(0, 10), rep(0, 10), rep(0, 10), rep(0, 10))
colnames(stats_table) <- c('n_missing1', 'n_missing2', 'n_missing3', 'n_missing4','stats','p')
for (iTime in 3:12){
  current_missing_counts <- missing_counts_wk10[,iTime]
  current_nonMissing_counts <- group_num_wk10[,2]-current_missing_counts
  current_table <- cbind(current_missing_counts, current_nonMissing_counts)
  results_chi <- chisq.test(current_table)
  stats_table[iTime-2,5] <- results_chi$statistic
  stats_table[iTime-2,6] <- results_chi$p.value
  stats_table[iTime-2,1:4] <- t(current_missing_counts)
}

# get the relative missing value number to the absolute number of subjects
missing_counts_wk10z <- missing_counts_wk10 %>%
  select(2:ncol(.)) %>%
  mutate_all(~./group_num_wk10$n)

# add a group and model label
missing_counts_wk10z <- cbind(group_num_wk10$wk10_model_groupF,missing_counts_wk10z)
names(missing_counts_wk10z)[1] <- "Group"

# coverting wide data to long data
missing_counts_wk10z_long <- missing_counts_wk10z %>%
  pivot_longer(
    cols = 2:ncol(.),
    names_to = "Weeks",
    values_to = "Missing_z"
  )
missing_counts_wk10z_long$Time <- factor(
          missing_counts_wk10z_long$Weeks, 
          levels = c("Y","W1PHQ9","W2PHQ9","W3PHQ9","W4PHQ9","W5PHQ9","W6PHQ9","W7PHQ9","W8PHQ9","W9PHQ9","W10PHQ9"), 
          labels = as.character(c(0:10))
          )

missing_counts_wk10z_long$Group <- factor(
                        missing_counts_wk10z_long$Group, 
                        levels = c(1,2,3,4),
                        labels = c("Rapid response, standard treatment length", "Rapid response, extended plateau", "Linear response, extended course", "Nonresponse"))

# Week 7 model
# count the absolute number of the missing value in each group
missing_counts_wk7 <- trajectory_data_clean %>%
  group_by(wk7_model_groupF) %>%
  select(25:35) %>%
  summarise_all(funs(sum(is.na(.))))

# get the number of participants in each group
group_num_wk7 <- count(trajectory_data_clean,wk7_model_groupF)
group_num_wk7

# get the relative missing value number to the absolute number of subjects
missing_counts_wk7z <- missing_counts_wk7 %>%
  select(2:ncol(.)) %>%
  mutate_all(~./group_num_wk7$n)

# add a group label
missing_counts_wk7z <- cbind(group_num_wk7$wk7_model_groupF,missing_counts_wk7z)
names(missing_counts_wk7z)[1] <- "Group"

# coverting wide data to long data
missing_counts_wk7z_long <- missing_counts_wk7z %>%
  pivot_longer(
    cols = 2:ncol(.),
    names_to = "Weeks",
    values_to = "Missing_z"
  )
missing_counts_wk7z_long$Time <- factor(
                  missing_counts_wk7z_long$Weeks, 
                  levels = c("Y","W1PHQ9","W2PHQ9","W3PHQ9","W4PHQ9","W5PHQ9","W6PHQ9","W7PHQ9","W8PHQ9","W9PHQ9","W10PHQ9"), 
                  labels = c(as.character(c(0:10)))
                  )
missing_counts_wk7z_long$Group <- factor(
                    missing_counts_wk7z_long$Group, 
                    levels = c(1,2,3,4),
                    labels = c("Rapid response", "Response, plateau", "Linear response", "Nonresponse")
                    )

```

## plot some supplementary figures

the bar plots and the line plot showing the ratios of missing values in each group

```{r}
# plot with ggplot2
# install.packages("ggplot2")
library(ggplot2)
# install.packages("devtools")
devtools::install_github("thomasp85/patchwork")
library(patchwork)

# plotting a bar plot showing each group's subject number
# week 10 model
ggplot(group_num_wk10, aes(x = wk10_model_groupF, y = n)) +
  geom_bar(stat = "identity") +
  xlab("Group") +
  ylab("Number of subjects") +
  ggtitle("Week 10 model") +
  theme_classic(base_size = 20)
ggsave("wk10model_subject_n.png", width = 10, height = 10, dpi = 600)

#week 7 model
ggplot(group_num_wk7, aes(x = wk7_model_groupF, y = n)) +
  geom_bar(stat = "identity") +
  xlab("Group") +
  ylab("Number of subjects") +
  ggtitle("Week 7 model") +
  theme_classic(base_size = 20)
ggsave("wk7model_subject_n.png", width = 10, height = 10, dpi = 600)

# plotting the line plots showing the missing percentage of each groups
# week 10 model
ggplot(missing_counts_wk10z_long, aes(x = Time, y = Missing_z, group = Group, color = Group)) +
  geom_line(size = 1.5) +
  labs(x = "Week", y = "Ratios of missing values", color = "Group") +
  theme_classic(base_size = 20) +
  scale_color_manual(values=c('#FF0060','#0079FF','#00DFA2','#FFE194'))+
  theme(legend.text = element_text(size = 16))

ggsave("ratio_missing_value.png", width = 13, height = 8, dpi = 600)

# week 7 model 

# p2 <- ggplot(missing_counts_wk7z_long, aes(x = Time, y = Missing_z, group = Group, color = Group)) + # geom_line(size = 1) + # labs(x = "Time", y = "Relative number of Missing Values", color = "Group") + # ggtitle("Relative number of missing values (Week 7 data)") + # theme_classic(base_size = 20) + # theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) # # ggsave("wk10model_relative_missing_value.png", width = 10, height = 8, dpi = 600)

```

## Plotting dot and line figures with shaded error bars, main results

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)

#+++++++++++++++++++++++++
# Function to calculate the mean and the standard deviation
  # for each group
#+++++++++++++++++++++++++
# data : a data frame
# varname : the name of a column containing the variable
  #to be summariezed
# groupnames : vector of column names to be used as
  # grouping variables
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}

# select data
data_dot_line <- trajectory_data_clean %>%
                select(c(
                        "Y","W1PHQ9", "W2PHQ9", "W3PHQ9", "W4PHQ9", "W5PHQ9",
                        "W6PHQ9", "W7PHQ9", "W8PHQ9", "W9PHQ9", "W10PHQ9",
                        "wk10_model_group"
                        )
                       )
# exclude group 4 from week 8-10
data_dot_line <- data_dot_line %>% 
                mutate(W8PHQ9 = ifelse(wk10_model_group == 4, NA, W8PHQ9),
                       W9PHQ9 = ifelse(wk10_model_group == 4, NA, W8PHQ9),
                       W10PHQ9 = ifelse(wk10_model_group == 4, NA, W10PHQ9)
                       )
names(data_dot_line)[1:11] <- as.character(c(0:10))
data_dtl_long <- data_dot_line %>%
              pivot_longer(
                cols = as.character(c(0:10)),
                names_to = "Time",
                values_to = "Value"
              )
df_summary <- data_summary(data_dtl_long, varname="Value", 
                    groupnames=c("Time", "wk10_model_group"))
names(df_summary)[2] <- "Group"
df_summary$Group <- as.factor(df_summary$Group)
df_summary$Group <- factor(df_summary$Group, levels = c("1", "2", "3", "4"), labels = c("Rapid response, standard treatment length", "Rapid response, extended plateau", "Linear response, extended course", "Nonresponse"))

# dot/line plot of week 10 model
ggplot(df_summary, aes(x= as.integer(Time), y=Value, group=Group, color=Group)) + 
  geom_line(size = 1) + xlab("Week") + ylab("PHQ-9 score") +
  geom_point(size = 3) +
  geom_ribbon(aes(ymin=Value-sd, ymax=Value+sd), fill = "grey70", alpha = .4, colour = NA) +
  scale_y_continuous(limits = c(-0.5, 30), breaks = seq(0, 30, by = 5)) +
  scale_x_continuous(limits = c(-0.5, 11), breaks = c(0:10)) +
  scale_color_manual(values=c('#FF0060','#0079FF','#00DFA2','#FFE194')) +
  theme_classic(base_size = 20) + 
  theme(legend.text = element_text(size = 12))

ggsave("wk10model_trajectory.png", width = 10, height = 5, dpi = 600)


# select data
data_dot_line <- trajectory_data_clean %>%
                select(c(
                        "Y","W1PHQ9", "W2PHQ9", "W3PHQ9", "W4PHQ9", "W5PHQ9",
                        "W6PHQ9", "W7PHQ9",
                        "wk7_model_group"
                        )
                       )
names(data_dot_line)[1:8] <- as.character(c(0:7))
data_dtl_long <- data_dot_line %>%
              pivot_longer(
                cols = as.character(c(0:7)),
                names_to = "Time",
                values_to = "Value"
              )
df_summary <- data_summary(data_dtl_long, varname="Value", 
                    groupnames=c("Time", "wk7_model_group"))
names(df_summary)[2] <- "Group"
df_summary$Group <- as.factor(df_summary$Group)
df_summary$Group <- factor(df_summary$Group, levels = c("1", "2", "3", "4"), labels = c("Rapid response, standard treatment length", "Rapid response, extended plateau", "Linear response, extended course", "Nonresponse"))

# dot/line plot of week 7 model
ggplot(df_summary, aes(x= as.integer(Time), y=Value, group=Group, color=Group)) + 
  geom_line(size = 1) + xlab("Week") + ylab("PHQ-9 score") +
  geom_point(size = 3) +
  geom_ribbon(aes(ymin=Value-sd, ymax=Value+sd), fill = "grey70", alpha = .4, colour = NA) +
  scale_y_continuous(limits = c(-0.5, 30), breaks = seq(0, 30, by = 5)) +
  scale_x_continuous(limits = c(-0.5, 8), breaks = c(0:7)) +
  scale_color_manual(values=c('#FF0060','#0079FF','#00DFA2','#FFE194')) +
  theme_classic(base_size = 20) + 
  theme(legend.text = element_text(size = 12))

ggsave("wk7model_trajectory.png", width = 10, height = 5, dpi = 600)
```
